---
title: "Intro to Data & Probability - Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---
<div id="intro">
The purpose of this project is to use an explortory data analysis to study several research questions which are to be formulated.
</div>

The dataset used in this project comes from the "Behavioral Risk Factor Surveillance System" (BRFSS) 2013 edition. Every year the US Center for Dease Control and Prevention (CDS) asks a conducts a significant questionnaire of health related questions  on a large sample of the US population resizind in all US states and territories. The  BRFSS 2013 edition  contains a dataset of over 491 thousands of residents with over 300 variables.  

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data


* * *

## Part 2: Research questions

**Research quesion 1:**

**Research quesion 2:**

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}

```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

